---
title: "ggplot-Week3-assignment"
output: html_document
author: Chrissy Spencer
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
```

# Exercise after Introduction to R and Tidyverse


```{r install tidyverse}

#install.packages("tidyverse")

```


Code chunk to load the tidyverse suite of packages

```{r load packages}

# use the library function to load dplyr, tidyr, palmerpenguins, and ggplot2

library(tidyverse)


```

Write a code chunk that:
-	Reads in socialMediaAddiction.csv to a variable called socialMedia
-	Produces a “glimpse” of the dataset


In this code along session, Palmer Penguins dataset where each observation is a penguin captured in a given year and each variable is a feature such as "species", "sex" and "body_mass_g". We will be exploring this data using tidyverse with a focus on [dplyr](http://dplyr.tidyverse.org/) for data wrangling, visualize it with [ggplot2](https://ggplot2.tidyverse.org/), and analyze it using the skills you've already developed!

Load your data:

```{r}
# Load the socialMediaAddiction dataset
# Dataset saved in the same place as .Rmd to load data properly

socialMediaAddiction <- read_csv("../data/socialMediaAddiction.csv")

# in RMD you don't have to indicate the file location but 
# use ../data/ to go up a folder and then into the data folder
# Check out a "glimpse" of the variables of your dataframe
glimpse(socialMediaAddiction)
socialMediaAddiction
# generates a tibble of socialMediaAddiction

```
Reviewed the dataset glimpse against the metadata provided 

# Data Visualization with ggplot2
Generate histograms for two continuous variables and create a code chunk that produces a histogram of each variable. 

Create a code chunk with opt-cmd-I
```{r histogram}
socialMediaAddiction %>%
  ggplot(mapping = aes(x = Avg_Daily_Usage_Hours)) +
  geom_histogram()

socialMediaAddiction %>%
  ggplot(mapping = aes(x = Mental_Health_Score)) +
  geom_histogram()

socialMediaAddiction %>%
  ggplot(mapping = aes(x = Sleep_Hours_Per_Night)) +
  geom_histogram()
```
QUESTION: is there a way to tab back and forth between graphs like in the Plots window? 
ANSWER: run full code chunk at once to get thumbnails for each histogram and tab between them.

# Describe the shape of each histogram
Avg_Daily_Usage_Hours is a normal-shaped distribution with some very jagged spots around 4.5 and 5 hours of daily social media usage (omg this is LOT of social media)

Mental_Health_Score is a discrete score with 1 = poor to 10 = excellent that might be skewed left with a tail to the right. No one selected low scores below 4, no one is a 10 and very few are 9s.

Sleep_Hours_Per_Night ranges from less than 4 to nearly 10 and looks roughly normal.


## Plotting your data

To plot your data using `ggplot2`, you specify three things:

* Your data (`penguins`)
* Your aesthetics (e.g. at minimium what's on the x-axis)
* Your layers (e.g. boxplot, scatterplot)


First you'll make a *histogram* to understand the distribution of penguin weights in your data. Histograms are used when you want to visualize the distribution of a single variable in your dataset.

```{r histogram}
# Plot a histogram of penguin weight
# using pipe syntax of %>% or 
# = also called "gets"
penguins %>%
  ggplot(mapping = aes(x = body_mass_g)) +
  geom_histogram()
# used two different connectors, pipe and pluses to add features
# this is dumb to have two different connectors but its too cumbersome to fix
# look for output in the rmd file!


```



Next you'll plot a boxplot of body mass by Sex to see how the body mass of male and female penguins differ. 


```{r boxplot}

# Plot boxplot of body mass by penguin Sex
penguins %>%
  ggplot(mapping = aes(x=sex, y=body_mass_g))+
  geom_boxplot()
  
# Plot boxplot of body mass by penguin species
penguins %>%
  ggplot(mapping = aes(x=species, y=body_mass_g))+
  geom_boxplot()

  

# Plot boxplot of body mass by penguin Sex separated by species
penguins %>%
  ggplot(mapping = aes(x=species, y=body_mass_g))+
  geom_boxplot()+
  facet_wrap(~sex)

penguins %>%
  ggplot(mapping = aes(x=species, y=body_mass_g, color = sex))+
  geom_boxplot()
  
penguins %>%
  ggplot(mapping = aes(x=species, y=body_mass_g, fill = sex))+
  geom_boxplot()


```

**What can you see here?** 

**Note:**

*  the `aes()` function is really a way to map the `aes`thetics of the plot to the variables in the data. See [Modern Dive](http://moderndive.com/3-viz.html#grammarofgraphics) by Chester Ismay & Albert Y. Kim for more.


Now it's time for some scatter plots. Is the bill length of a penguin associated with penguin weight? What if we separate our data by species?


```{r scatterplot}

# Scatterplot of bill length and weight
penguins %>%
  ggplot(mapping = aes(x=bill_length_mm,y=body_mass_g))+
  geom_point()+
  geom_smooth(method = lm)
#smooth generates a lowess fcn, add method to make it streight line



# Scatterplot of bill length faceted by penguin species



```

**What can you see here?** 

**Note:**

* You can read your `ggplot2` code like a sentence, just like your `dplyr` code above: "We take the data as penguins and we map bill length to the x axis and weight to the y axis, adding points on as the layer of the plot."


**Recap:**



## Exercise (Time-Permitting)

We will now use some of the skills we've just learnt to answer a our own questions! 

As a group: 

1.  Develop a question that can be answered with the `penguins` dataset. 
2.  Talk through how the data needs to be manipulated to get the appropriate answer. Would plotting the data be useful? Which verbs will you use and in what order? 
3.  Write the code for the analysis. Designate a computer that will have the completed code. If you are not actively typing, you can help your group by thinking about what should come next or proof reading what has already been typed. Be sure to test your code often. 


```{r}

# Write your analysis here




```
